\chapter{Project Specification}

\section{Aims}

The aim of SystemNaim is to create a HLS tool which allows users to target a multi-FPGA system. Most modern HLS tools, which will be touched on later, require the user to manually specify the split of the system and implement their own communication protocol. This project intends to alleviate the effort that must be put in by the user by automating much of the process.

\section{Motivation}

Before we go into detail about the specifications of the tool, we must first understand the necessity of it. This section explores why FPGAs, HLS tools and multi-FPGA systems are each important areas of research and gives insight into the motivation for this project.

\subsection{FPGAs}

FPGAs(Field-Programmable Gate Arrays) have been around for since the 1980s and have served various purposes over the years\cite{fpga-history}. They boast a faster time-to-market than ASICs whilst still providing a performance boost when compared to software running on a GPPs(General purpose processor). These features, amongst others, make FPGAs favourable for a variety of uses including:
\begin{itemize}
    \item Prototyping.
    \item Hardware Emulation.
    \item Data processing for data centres.
    \item High-frequency trading.
 \end{itemize}

In more technical terms an FPGA is a large collection of LE(Logic Elements) which can programmed using an HDL(Hardware Description Language) such as Verilog. Each of these LE can only perform a small amount of logic such as implementing an AND/OR gate but an average FPGA comes with tens of thousands of them allowing for large amounts of logic to be strung together. The important fact is that, once programmed, an FPGA becomes a piece of hardware dedicated to performing a fixed operation and thus, if programmed correctly, gives lower latencies and higher-throughputs alongside lower power requirements than the same operation being conducted in software on a GPP. In summary, if a user is looking for a way to decrease the latency, or increase the throughput, of a rapidly changing system, an FPGA would definitely be a strong contender for a solution. 

\subsection{HLS Tools}

Notice however, that in the section above it was explicitly mentioned that benefits of using an FPGA only came if the device in question is programmed correctly. Now correctness in this context refers to multiple things. Firstly, the operation must be suited to an FPGA. This can include data access patterns or whether the application has a large amount of parallelism the programmer can exploit, but the key takeaway is that not all programs are suited to being executed on an FPGA with some performing better on modern GPPs, and it is thus up to the designer/programmer to make this judgement call.

In this report, however, we are more concerned about another aspect of correctness; the difficulty of writing HDL code. FPGA development is notoriously difficult due to the lack of individuals with the skills to write HDL code specifically targeted at FPGAs. The cause of this issue is two-fold, one is that FPGA usage isn't very well known and the second being the paradigm shift between programming in languages such as: C++ or python, and programming in a HDL like Verilog. An entire report could be written to explain the differences between hardware and software programming, but the main point is that it is very difficult to shift between the two paradigms which results in fewer people who have the skills to exploit the benefits of an FPGA effectively. To help solve this problem, HLS(High-Level Synthesis) tools have been actively researched over the last 40 years~\cite{5209959}, and now we have a variety of products such as Vivado HLS from Xilinx, Intel® High Level Synthesis Compiler from Intel and LegUp from Microchip.

These tools allow users to write code in C++/C, which is then transpiled to Verilog for use on an FPGA. In order to create optimised code these tools include configuration options and extra syntax building upon the language they are based in, thus allowing the programmer more control over how their code is turned into hardware, this addition does mean there is some learning curve to using these tools however it is much less than going from C++ to Verilog. 

Overall these tools allows for a much easier transfer of skills from normal programming to hardware design, paving the way for a larger part of industry and academia to gain the potential benefits of using an FPGA. Therefore, research into making these tools more accessible and on par, performance wise, with pure Verilog design is very appealing when one considers the rising usage FPGAs in variety of fields\cite{fpga-market}.

\subsection{Multi-FPGA Communication}

One of the major limitations of any hardware implementation is the resources available to a programmer on the device. Examples of such resources are:

\begin{itemize}
    \item \textbf{BRAMs}: For on-chip storage.
    \item \textbf{DRAM Bandwidth}: For accessing off-chip memory.
    \item \textbf{DSPs}: For floating-point arithmetic.
    \item \textbf{LUTs}: For logic processing.
\end{itemize}

If the programmer finds themselves having exhausted the resources on the board but still in need of more, they have a few options that they can explore. The first and usually the most practical is to re-structure the design. This may include adding pipeline stages to remove timing violations, moving to data storage off-chip to increase on-chip memory space or reusing modules in multiple parts of the design to save on LUTs. All optimisations while viable do have downsides and considerations the programmer must take in to account before proceeding. The next option would be to get a device with more resources available, while ideal this option may not be possible either because the new device would be too expensive or more likely the programmer has been given a specification and a part of it is to use this specific device. The third option and the one of interest to this report is to use multiple FPGAs in parallel to compute different parts of the system giving the programmer access to more resources in doing so.

Utilising a multi-FPGA system to overcome the limitations of a single FPGA has been the focal point for many research fronts including: DNN Inference\cite{10.1145/3358192}, Energy Conversion System Simulation\cite{8822485}, Large Graph Processing\cite{10.1145/3020078.3021739}, and CNN Acceleration \cite{10.1145/3337821.3337846} and thus proves there is benefit to allowing more people access to such systems. There are of course drawbacks for this method such as the overhead communication cost between the two FPGAs and those will be explored later in the report.

\section{Deliverables}

This project consists mainly of two deliverables, one is the HLS tool, SystemNaim, which is entirely based in software and the other is an analytical report of the tool itself. This report will evaluate the efficacy of the SystemNaim as well as indicate areas where further development would be beneficial.

\subsection{SystemNaim}

As previously stated, the aim of SystemNaim is to create a HLS tool which allows users to target a multi-FPGA system. However, we must acknowledge that this being an FYP there a time constraint which will limit the amount of work possible. Therefore, some limitations have been placed on the scope of SystemNaim in order to ensure that a working tool can be produced. They are as follows:

\begin{itemize}
    \item The input will be a single source file following a subset of ANSI C90. The limitations of the language are:
    \begin{itemize}
        \item Only integer types will be allowed.
        \item Only local arrays will be allowed.
        \item Structs \& enums will not be allowed.
    \end{itemize}
    \item Only two FPGAs will communicate with each other.
\end{itemize}

It should be noted that these limitations are subject to change and certain design choices will introduce more. Also, if work on the base tool is completed sooner than expected, some of these limitations will be removed and work will be done to increase scope of the tool. If this is the case then the two expansions that would be of most interest are the allowing of passing arrays between functions and more than two FPGAs communicating.

The output of SystemNaim will be a collection of Verilog HDL files which perform the same computation as the provided input file and with each FPGA receiving its own top file. The target devices have yet to be decided, however, it is likely that regardless of this choice the output files will simply be accompanied by a guide on how to implement the design using the currently available synthesis tools. It is also yet to be decided whether pre-existing libraries for base functionality, such as arithmetic and memory access, will be used. This choice will be made after the target device is chosen and depends upon the difficulty of incorporating these libraries into the tool.

To prove the tool works, a minimum of two FPGAs will be used for test purposes. While these are not part of the deliverable, it is worth noting that there is an intrinsic hardware component to this project.


\subsection{Analytical Report/ Evaluation Plan}

The analysis report serves to test whether the features implemented in SystemNaim offers any benefit to the user. While all projects need a report such as this, it is much more important to this FYP as its entire purpose is to improve quantifiable metrics. To be specific, time-to-market and performance are the key metrics this tool should improve upon. It should be easier for a user to create a multi-FPGA system using SystemNaim than existing tools, and they should be able to observer a measurable speed up when doing so.

\subsubsection{User Difficulty}
For the time-to-market metric we intend to compare against other available HLS tools as well as writing a system in pure Verilog. Not only will the time-taken be considered but also the amount of extra research a user must do in order to achieve their desired design. Ideally, this evaluation will be done by asking a sample of individuals to implement a multi-FPGA system using a variety of tools.

\subsubsection{Performance}
On the other hand, the performance metric has a few more complications that we must consider. 

Firstly, performance is a vague term, and we must define more specific metrics in order to measure this value. There are two main metrics that shall be used in this report: 

\begin{itemize}
    \item \textbf{Latency}: The amount of time it takes for a system to compute it's intended function. Measured either in milliseconds or cycles.
    \item \textbf{Resource Usage}: The amount of resources a system takes up on an FPGA. Measures in number of LUTs, BRAMs and DSPs.
\end{itemize}

These metrics can be further broken down, however for the purpose of this project these will tell us all the information we need\footnote{Throughput may also be seen as important however with loop pipe-lining likely not being implemented it serves little purpose.}.

Tools such as Vivado HLS or LegUp have spent years in development and thus comparing the latency of a design produced by a product of their calibre versus something made in less than 6 months would not benefit any evaluation of the specific issue SystemNaim aims to solve, creating multi-FPGA systems. If we were to compare SystemNaim to an industry tool, we wouldn't be able to see the benefit of using multiple FPGAs to increase the performance of a system. Rather, we would merely see how much better optimised the Verilog that the industry tools produce is, over SystemNaim's output Verilog.

Therefore, the methodology we intend to use to measure performance is to compare the latency and resource usage of a single FPGA system created using SystemNaim and a multi-FPGA system created from the same tool. This way we can identify the performance increase/decrease specifically attributed to the tool's method of implementing multi-FPGA systems. 

The conclusion that we hope to draw from this, is that if this method of automating the connection between multiple FPGAs is beneficial then it is possible to expand this methodology to other tools. So even if the core HLS aspect of SystemNaim is subpar, the process by which it creates multi-FPGA systems adds value to the field of research.






